---
title: "Driver"
output:
  output: pdf_document
---

<!-- General To-Do -->

<!-- * Merge the Datasets! -->
<!-- * Data Cleaning -->
<!--     * Some rows in Seasons_Stats have nothing in them but the index--remove -->


<!-- Statistical Analysis Ideas: -->

<!-- * Multivariate Analysis -->
<!--     * Cluster Analysis -->
<!--     * PCA -->
<!-- * Data Visualization -->
    <!-- * Make a creative, "original" plot--perhaps use grid? -->
    <!-- * Make 3-d plot interactive! -->
    <!-- * trellis plots to combine the variables -->
    <!--     * Cleveland Caveliers vs. other teams, and other variables  -->
<!-- * Inference -->
<!--     * Missing Data -->
<!--     * MLR plus model selection -->
<!--     * Correlations!!!!!!!!!! Make plots, do PCA -->
<!-- * Prediction -->
<!--     * Random Forest -->
<!--     * Cross-Validation -->
<!--     * Model Selection -->
<!--         * Elastic-Net Regularization -->
<!-- * Other relevant stuff Tianyi cited -->



```{r}
library(ggplot2)
library(scatterplot3d)
library(glmnet)
```

```{r}
sal_1718 <- read.csv("NBA_season1718_salary.csv")
sal_1819 <- read.csv("NBA_season1819_salary.csv")
player_data <- read.csv("player_data.csv")
Players <- read.csv("Players.csv")
season <- read.csv("Seasons_Stats.csv")
```

# Merge the Data

## Salary with season, for years 

```{r}
# getting rid of index in both datasets
# getting rid of two blank columns in season data set, as well as Year, since it'll all be 2017
sal_season <- merge(sal_1718[,c(-1)], season[season$Year == 2017,-c(1, 2, 22, 27)], by = c("Player", "Tm"))

write.csv(sal_season, file = "sal_season.csv")
```
<!-- After merging, do 3d scatterplot of salary vs. other stuff -->

# Linear Model

## Correlation Analysis

```{r}
sal_season_mat <- model.matrix(~ ., sal_season[,c(-1, -3)])[,-1] 
# inner negatives: remove name and salary (dependent variable). Outer negative: remove intercept

sal <- sal_season$season17_18[rowSums(is.na(sal_season)) == 0]

R <- cor(sal_season_mat, method = "spearman")
```

```{r, include = FALSE}
# Which correlations are > .50 (in absolute value) but not 1?
which(abs(R) > .50 & R != 1, arr.ind = TRUE)

R[which(abs(R) > .50 & R != 1, arr.ind = TRUE)]
```

Too big of a correlation matrix to display. 

## Graphics

Scatterplots between salary and arbitrarily chosen predictors
```{r}
plot(sal_season$Age, sal_season$season17_18)

plot(sal_season$FTr, sal_season$season17_18)

plot(sal_season$STL., sal_season$season17_18)

plot(sal_season$WS, sal_season$season17_18)

plot(sal_season$FGA, sal_season$season17_18)

plot(sal_season$X2P., sal_season$season17_18)

plot(sal_season$TRB, sal_season$season17_18)

plot(sal_season$PF, sal_season$season17_18)
```

log-transforming salary
```{r}
plot(sal_season$Age, log(sal_season$season17_18))

plot(sal_season$FTr, log(sal_season$season17_18))

plot(sal_season$STL., log(sal_season$season17_18))

plot(sal_season$WS, log(sal_season$season17_18))

plot(sal_season$FGA, log(sal_season$season17_18))

plot(sal_season$X2P., log(sal_season$season17_18))

plot(sal_season$TRB, log(sal_season$season17_18))

plot(sal_season$PF, log(sal_season$season17_18))
```

Pairwise Plots
```{r}
pairs(~ ., data = sal_season[,-c(1, 2, 4, 9:50)])

pairs(~ ., data = sal_season[,-c(1:8, 15:50)])

# pairs(~ ., data = sal_season[,-c(1:8, 15:50)])
# 
# pairs(~ ., data = sal_season[,-c(1, 2, 3, 4, 5:26)])
# 
# pairs(~ ., data = sal_season[,-c(1, 2, 4, 27:50)])
# 
# pairs(~ ., data = sal_season[,-c(1, 2, 3, 4, 5:26)])
# 
# pairs(~ ., data = sal_season[,-c(1, 2, 3, 4, 5:26)])
# 
# pairs(~ ., data = sal_season[,-c(1, 2, 3, 4, 5:26)])
```

## Model

Ersan Ilyasova appears twice in the data
```{r}
sal_lm <- lm(season17_18 ~ ., data = sal_season[,-1])

# taking out the Player name as a predictor

# sal_lm_log <- lm(log(season17_18) ~ ., data = sal_season[,-1]) 
```
<!-- elastic-net regularize this -->

```{r}
summary(sal_lm)
```

```{r}
plot(sal_lm$fitted.values, rstandard(sal_lm), main = "Residuals vs. Fitted Values", xlab = "Fitted Values", ylab = "Standardized Residuals")
abline(h = 0, col = "red")

# plot(sal_lm_log$fitted.values, rstandard(sal_lm_log), main = "Residuals vs. Fitted Values", xlab = "Fitted Values", ylab = "Standardized Residuals")
# abline(h = 0, col = "red")
```
Looks screwy. Let's log-transform the data--looking on kaggle, other ppl didn't do it

## Elastic-Net Regularization

Finding the optimal alpha
```{r}
set.seed(1262019)
n <- nrow(sal_season_mat)
# Determining a random foldid
foldid <- sample(1:10, # default is 10 folds
                 size = n,
                 replace = TRUE
                 )

# Sequence of alphas to test
alphas <- seq(0, 1, by = .05)

devs <- rep(0, length(alphas))

for (i in 1:length(alphas)) {
cv <- cv.glmnet(sal_season_mat, sal, foldid = foldid, alpha = alphas[i], keep = TRUE)
  devs[i] <- min(cv$cvm)
}

devs
```
alpha of 1 is the best--just use alpha = .95 to strike a balance b/w ridge and lasso

```{r}
set.seed(123)

cv_lm <- cv.glmnet(sal_season_mat, sal, alpha = 1)
```

```{r}
plot(cv_lm$glmnet.fit, label = TRUE)
plot(cv_lm)


coef(cv_lm, s = cv_lm$lambda.min)
```

## Constructing LM from the elastic-net regularization 

```{r}
TmBOS <- ifelse(sal_season$Tm == "BOS", 1, 0)
TmBRK <- ifelse(sal_season$Tm == "BRK", 1, 0)
TmCHI <- ifelse(sal_season$Tm == "CHI", 1, 0)
TmCLE <- ifelse(sal_season$Tm == "CLE", 1, 0)
TmDAL <- ifelse(sal_season$Tm == "DAL", 1, 0)
TmGSW <- ifelse(sal_season$Tm == "GSW", 1, 0)
TmHOU <- ifelse(sal_season$Tm == "HOU", 1, 0)
TmIND <- ifelse(sal_season$Tm == "IND", 1, 0)
TmLAC <- ifelse(sal_season$Tm == "LAC", 1, 0)
TmLAL <- ifelse(sal_season$Tm == "LAL", 1, 0)
TmMEM <- ifelse(sal_season$Tm == "MEM", 1, 0)
TmMIA <- ifelse(sal_season$Tm == "MIA", 1, 0)
TmMIN <- ifelse(sal_season$Tm == "MIN", 1, 0)
TmNOP <- ifelse(sal_season$Tm == "NOP", 1, 0)
TmOKC <- ifelse(sal_season$Tm == "OKC", 1, 0)
TmORL <- ifelse(sal_season$Tm == "ORL", 1, 0)
TmPHI <- ifelse(sal_season$Tm == "PHI", 1, 0)
TmPOR <- ifelse(sal_season$Tm == "POR", 1, 0)
TmSAC <- ifelse(sal_season$Tm == "SAC", 1, 0)
TmSAS <- ifelse(sal_season$Tm == "SAS", 1, 0)
TmTOR <- ifelse(sal_season$Tm == "TOR", 1, 0)
TmUTA <- ifelse(sal_season$Tm == "UTA", 1, 0)
TmWAS <- ifelse(sal_season$Tm == "WAS", 1, 0)

PosPG <- ifelse(sal_season$Pos == "PG", 1, 0)
PosSF <- ifelse(sal_season$Pos == "SF", 1, 0)
PosSG <- ifelse(sal_season$Pos == "SG", 1, 0)

sal_lm_en <- lm(season17_18 ~ TmBOS + TmBRK + TmCHI + TmCLE + TmDAL + TmGSW + TmHOU + TmIND + TmLAC + TmLAL + TmMEM + TmMIA + TmMIN + TmNOP + TmOKC + TmORL + TmPHI + TmPOR + TmSAC + TmSAS + TmTOR + TmTOR + TmUTA + TmWAS + PosPG + PosSF + PosSG + Age + G + GS + TS. + X3PAr + STL. + USG. + DWS + WS + X3P + X3P. + X2P. + FT. + AST + PF, data = sal_season)
```

```{r}
summary(sal_lm_en)
```

```{r}
plot(sal_lm_en$fitted.values, rstandard(sal_lm_en), main = "Residuals vs. Fitted Values", xlab = "Fitted Values", ylab = "Standardized Residuals")
abline(h = 0, col = "red")
```


## Plots based on model

```{r}
plot(sal_season$Age, sal_season$season17_18, col = 6)
plot(sal_season$PTS, sal_season$season17_18, col = 6)
```

### 3d scatterplots

```{r}
s3d_1 <- scatterplot3d(x = season$Age[season$Pos == "C"], y = season$MP[season$Pos == "C"], z = season$TS.[season$Pos == "C"], type = "p", color = 1)

positions <- levels(season$Pos)[3:24]
index <- 2

for (pos in positions) {
  s3d_1$points3d(x = season$Age[season$Pos == pos], y = season$MP[season$Pos == pos], z = season$TS.[season$Pos == pos], type = "p", col = index)
  index = index + 1
}
```


```{r}
s3d_1 <- scatterplot3d(x = sal_season$Age[sal_season$Pos == "C"], y = sal_season$MP[sal_season$Pos == "C"], z = sal_season$season17_18[sal_season$Pos == "C"], type = "p", color = 1, xlab = "Age", ylab = "Minutes Played", zlab = "Salary")

positions <- levels(sal_season$Pos)[3:24]
index <- 2

for (pos in positions) {
  s3d_1$points3d(x = sal_season$Age[sal_season$Pos == pos], y = sal_season$MP[sal_season$Pos == pos], z = sal_season$season17_18[sal_season$Pos == pos], type = "p", col = index)
  index = index + 1
}
```



Factor variable: Team 

```{r}
s3d_1 <- scatterplot3d(x = sal_season$Age[sal_season$Tm == "ORL"], y = sal_season$PTS[sal_season$Tm == "ORL"], z = sal_season$season17_18[sal_season$Tm == "ORL"], type = "p", color = 1, xlab = "Age", ylab = "PTS", zlab = "Salary")

teams <- levels(sal_season$Tm)[2:24]
index <- 2

for (tm in teams) {
  s3d_1$points3d(x = sal_season$Age[sal_season$Tm == tm], y = sal_season$PTS[sal_season$Tm == tm], z = sal_season$season17_18[sal_season$Tm == tm], type = "p", col = index)
  index = index + 1
}
```


# Exploration of the salary files

## Exploration of sal_1718

```{r}
hist(sal_1718$season17_18)

hist(sal_1819$seson18_19)
```

# Exploratory Data Analysis of Seasons_Stats

```{r}
sub_idx <- sample(nrow(season), size = 1000)

sub_season <- season[sub_idx,]
```

<!-- So the year only goes up to 2017. Which salary season does it correspond to? Just use the 18-19 season  -->
```{r}
summary(season)

str(season)
```

## 3d Scatterplot of subseason

```{r}
s3d_1 <- scatterplot3d(x = sub_season$Age[sub_season$Pos == "C"], y = sub_season$MP[sub_season$Pos == "C"], z = sub_season$TS.[sub_season$Pos == "C"], type = "p", color = 1)

positions <- levels(sub_season$Pos)[3:24]
index <- 2

for (pos in positions) {
  s3d_1$points3d(x = sub_season$Age[sub_season$Pos == pos], y = sub_season$MP[sub_season$Pos == pos], z = sub_season$TS.[sub_season$Pos == pos], type = "p", col = index)
  index = index + 1
}


```

## Principal Components Analysis

```{r}
pr.out <- prcomp(sal_season_mat[,-which(colSums(sal_season_mat) == 0)], scale = TRUE) 
# remember, sal_season_mat doesn't have the dependent variable. 
# therefore, the principal components can be used to predict the dependent variable
# -which... takes out rows with no zeros
```

```{r}
pc_dat <- pr.out$x


```

```{r}
sal_st <- (sal - mean(sal)) / sd(sal)

s3d_pc <- scatterplot3d(x = pc_dat[sal_season_mat[,"TmBOS"] == 1,1], y = pc_dat[sal_season_mat[,"TmBOS"] == 1,2], z = sal_st[sal_season_mat[,"TmBOS"] == 1], type = "p", color = 1, xlab = "PC1", ylab = "PC2", zlab = "Standardized Salary", z.ticklabs = NULL)

teams <- colnames(sal_season_mat)[2:24]
index <- 2

for (tm in teams) {
  s3d_pc$points3d(x = pc_dat[sal_season_mat[,tm] == 1,1], y = pc_dat[sal_season_mat[,tm] == 1,2], z = sal_st[sal_season_mat[,tm] == 1], col = index)
  index = index + 1
}
```

<!-- ```{r} -->
<!-- pr.out$rotation -->
<!-- ``` -->



## Model

```{r}
sal_lm_pc <- lm(sal ~ ., data = as.data.frame(pc_dat))

# taking out the Player name as a predictor

# sal_lm_log <- lm(log(season17_18) ~ ., data = sal_season[,-1]) 
```
<!-- elastic-net regularize this -->

```{r}
summary(sal_lm_pc)
```

```{r}
plot(sal_lm_pc$fitted.values, rstandard(sal_lm_pc), main = "Residuals vs. Fitted Values", xlab = "Fitted Values", ylab = "Standardized Residuals")
abline(h = 0, col = "red")

# plot(sal_lm_log$fitted.values, rstandard(sal_lm_log), main = "Residuals vs. Fitted Values", xlab = "Fitted Values", ylab = "Standardized Residuals")
# abline(h = 0, col = "red")
```

## Elastic-Net Regularization

Finding the optimal alpha
```{r}
set.seed(1262019)
n <- nrow(sal_season_mat)
# Determining a random foldid
foldid <- sample(1:10, # default is 10 folds
                 size = n,
                 replace = TRUE
                 )

# Sequence of alphas to test
alphas <- seq(0, 1, by = .05)

devs <- rep(0, length(alphas))

for (i in 1:length(alphas)) {
cv <- cv.glmnet(pc_dat, sal, foldid = foldid, alpha = alphas[i], keep = TRUE)
  devs[i] <- min(cv$cvm)
}

devs
```
alpha of 1 is the best

```{r}
set.seed(123)

cv_lm <- cv.glmnet(pc_dat, sal, alpha = 1)
```

```{r}
plot(cv_lm$glmnet.fit, label = TRUE)
plot(cv_lm)


coef(cv_lm, s = cv_lm$lambda.min)
```

## Constructing LM from the elastic-net regularization 

```{r}
sal_lm_en_pc <- lm(sal ~ PC1 + PC2 + PC6 + PC7 + PC8 + PC10 + PC13 + PC19 + PC22 + PC27 + PC28 + PC29 + PC31 + PC32 + PC34 + PC37+ PC41 + PC43 + PC44 + PC45 + PC46 + PC51 + PC52 + PC53 + PC57 + PC60 + PC61 + PC63 + PC64 + PC66+ PC70 + PC72 + PC73 + PC74+ PC78 + PC79, data = as.data.frame(pc_dat))
```

```{r}
summary(sal_lm_en_pc)
```

```{r}
plot(sal_lm_en_pc$fitted.values, rstandard(sal_lm_en_pc), main = "Residuals vs. Fitted Values", xlab = "Fitted Values", ylab = "Standardized Residuals")
abline(h = 0, col = "red")
```



## Cluster Analysis
